Bayesian Sparsification of $\cplx$-valued networks

### 2019-11-12 paper/paper.tex
in-text and parenthetical citations

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

@@

Deep neural networks are an integral part of machine learning and data science toolset
for practical data-driven problem solving. Despite recent advances in generalization
and optimization theory specific to deep networks, deploying in actual embedded hardware
remains a challenge. Solving the storage and compute constraints encourages development
of model compression and sparsification methods, that focus on favourable tradeoff between
performance and size. Most popular techniques include weight pruning \cite{citation_needed},
model quantization from IEEE754 floating point to integer-based arithmetic \cite{citation_needed},
and dropout. The key assumption is that the more explicitly zero parameters there is,
the less computations is needed.

Bayesian Inference is a principled framework of reasoning about uncertainty. The core principle
is updating prior beliefs in accordance with the likelihood of empirical observations into
posterior belief. Under suitable modelling and prior assumptions, Bayesian methods can be
used towards model sparsification, \cite{kingma_variational_2015,molchanov_variational_2017},
and \cite{kharitonov_variational_2018}.

% complex nets and related stuff
data that is naturally complex-valued
studies have shown a potentially richer representation of complex nets
\cite{hirose2003,fijulamin2009}
% [3] Akira Hirose, Complex-valued neural networks: theories and applications, vol. 5, World Scientific, 2003.
% [4] Md Faijul Amin and Kazuyuki Murase, “Single-layered complex-valued neural network for real-valued classification problems,” Neurocomputing, vol. 72, no. 4-6, pp. 945–955, 2009.

@@

\cite{yang_complex_2019} develop complex transformer, with $\cplx$-attention and complex
encoder-decoder, and apply it to music transcription and wireless signal classification.

@@

papers related to complex-valued distributions \cite{pav_moments_2015,taubock_complex-valued_2012},
and \cite{karseras_caution:_2014}

Application of complex-valued networks is not new, \cite{citation_needed}. More recently
they were applied to music transcription and speech recognition \cite{trabelsi_deep_2017},
and \cite{yang_complex_2019}. The key motivation behind adoption of $\cplx$-networks is
pervasiveness of $\cplx$-valued data in digital signal processing \cite{citation_needed},
acoustic and radio signal analysis \cite{citation_needed}, and quantum systems modelling
\cite{citation_needed}, etc. All these applications require compliance with algebraic
operations in the field of complex numbers.

@@

The posterior itself is computationally intractable, save for the relatively simple cases
\cite{citation_needed}, and therefore exact inference is traded for tractability and scalability
of Variational Inference. Instead of deriving $p(\omega \mid D)$ using the rule, this approach
recasts posterior inference into variational optimization problem: finding $q$ that is close
to $p(\omega \mid D)$ in terms of a proximity score $\rho$
\begin{equation}  \label{eq:variational-progam}
  q_*
    \in \arg \min_{\theta} \rho\bigl(
      q_\theta(\omega), p(\omega \mid D)
    \bigr)
    \,,
\end{equation}
over a tractable parametric family of distributions on the parameter space $
  \mathcal{Q} = \{Q_\theta(d\omega) \colon \theta \in \Theta\}
$, $\theta$ -- generic variational parameter. The class $\mathcal{Q}$ is picked so that
its members have tractable densities $
  q_\theta(\omega) d\omega
$, can be sampled from and possess tractable $\log$-derivatives $
  \omega \mapsto \nabla_\theta \log q_\theta(\omega)
$, \cite{williams_simple_1992}, or can be represented as differentiable transformations
of non-parametric noise, \cite{kingma_auto-encoding_2014,figurnov_implicit_2019}.
% (reparameterization trick, pathwise gradieent) $q_{\theta}(d\omega)$ is a push-forward
% of $p(d\varepsilon)$ by a differentiable map $g(\varepsilon; \theta)$.
% % http://stillbreeze.github.io/REINFORCE-vs-Reparameterization-trick/#fn:1

Kullback-Leibler divergence of density $q$ from $p$ is
\begin{equation}  \label{eq:kl-div-def}
  KL(q \| p)
    % = \int \frac{dQ}{dP} \log{\frac{dQ}{dP}} P(d\omega)
    % = \int \frac{q(\omega)}{p(\omega)} \log{\frac{q(\omega)}{p(\omega)}} p(\omega) d\omega
    % = \mathbb{E}_{\omega \sim q}
    %   \log\frac{q(\omega)}{p(\omega)}
    = \mathbb{E}_{\omega \sim q}
      \log{q(\omega)}
    - \mathbb{E}_{\omega \sim q}
      \log{p(\omega)}
    \,,
\end{equation}
% and, despite it not being a metric and requiring distributions with nested support,
and is the natural choice for $\rho$, the main reason being that it satisfies the following
identity for well-behaved densities:
\begin{equation}  \label{eq:kl-div-master}
  \log p(D)
    % = \mathbb{E}_{q(\omega)} \log \frac{p(D, \omega)}{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{q(\omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    % - \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{\pi(\omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
    = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    - KL(q(\omega) \| \pi(\omega))
    + KL(q(\omega) \| p(\omega \mid D))
    \,.
\end{equation}
This identity yields an equivalent problem for \eqref{eq:variational-progam}:
maximizing the \textit{evidence lower bound} (ELBO)
\begin{equation}  \label{eq:elbo_general}
\begin{aligned}
  & \underset{\theta, \phi, \lambda}{\text{maximize}}
    & \mathcal{L}(\theta; \phi, \lambda)
      & = \mathbb{E}_{\omega \sim q_{\theta}}
          \log p_{\phi}(D \mid \omega)
        - KL(q_{\theta}(\omega) \| \pi_{\lambda}(\omega))
        % \mathbb{E}_{\omega \sim q_{\theta}}
          % \log\frac{q_\theta(\omega)}{\pi_{\lambda}(\omega)}
  % \\
  % & &
  %     & = \mathbb{E}_{\omega \sim q_{\theta}}
  %         \log{p_{\phi}(D \mid \omega) \pi_{\lambda}(\omega)}
  %       + \mathbb{H}(q_{\theta})
  \,,
\end{aligned}
\end{equation}
where the prior and likelihood are allowed to depend on explicit parameters. The most
commonly used algorithm to optimize \eqref{eq:elbo_general} is EM-algorithm. Other
variational objectives are possible, provided $p(\omega \mid D)$ is probed only through
$\log p(D \mid \omega)$ and $\log \pi(\omega)$, \cite{ranganath_operator_2018}.

\textit{Stochastic Gradient Variational Bayes} (SGVB), proposed by \cite{kingma_auto-encoding_2014},
is a differentiable unbiased Monte-Carlo estimator of \eqref{eq:elbo_general}, which makes
it possible to employ stochastic gradient-based optimization methods. If $
  \omega \sim q_{\theta}
$ is equivalent in distribution to $
  \omega = g(\varepsilon; \theta)
$ for some non-parametric random variable $
    \varepsilon \sim p_\varepsilon
$ and differentiable $
  g(\varepsilon; \theta)
$, then the SGVB is
\begin{equation}  \label{eq:sgvb}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta}(\omega) \| \pi_{\lambda}(\omega))
      % + \frac{N}{\lvert B \rvert} \sum_{i\in B}
        % \log p(z_i \mid \omega_i)b
      + \frac1{L} \sum_{l=1}^L
        \log p_{\phi}(D \mid \omega_{il})
          \Big\vert_{\omega_{l} = g(\varepsilon_{l}; \theta)}
      % =
      % - KL(q_{\theta}(\omega) \| \pi(\omega))
      % + N \mathbb{E}_{B \sim \{D\}_M \otimes q_\theta^M}
        % \hat{\mathbb{E}}_{z,\omega \sim B}
          % \log p(z \mid \omega)
    \,, 
\end{equation}
where $
  (\varepsilon_{l})_{l=1}^L
$ is an iid sample from $p_\varepsilon$. Intractable KL-divergence terms can be replaced
by differentiable unbiased Monte-Carlo estimators, \cite{kingma_auto-encoding_2014}.
%
For the purposes of this study, we assume that the observed data is independent, and
use minibatch version of \eqref{eq:sgvb} with $L=1$:
\begin{equation}  \label{eq:elbo}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta}(\omega) \| \pi_{\lambda}(\omega))
      + \frac{N}{M} \sum_{i=1}^M
        \log p_{\phi}(z_i \mid g(\varepsilon_i; \theta))
    \,,
\end{equation}
for a random subsample $(z_i)_{i=1}^M$ from $D$ and iid $
  (\varepsilon_i)_{i=1}^M \sim p_\varepsilon
$.
%
\todo{DSVI}

@@

With special family of posterior approximation, variational inference can be used as
a regularization method, \cite{kingma_variational_2015}, and as model sparsification
technique, \cite{molchanov_variational_2017}. The paper \cite{kingma_variational_2015}
considers Dropout \cite{hinton_improving_2012}, DropConnect, \cite{wan_regularization_2013},
and Gaussian dropout, \cite{srivastava_dropout_2014,wang_fast_2013} through the lens
of Bayesian inference methods and proposes \textit{Variational Dropout}. It argues that
the multiplicative noise introduced by these methods induces a distribution equivalent to
a fully factorized variational posterior of the form $
  q_\theta(\omega) = \prod_j q_{\theta}(\omega_j)
$, where $q_{\theta}(\omega_j)$ is $\omega_j = \mu_j \xi_j$ with $
  \xi_j \sim p_\theta(\xi_j)
$ iid from some $p_\theta(\xi)$.
%
% Binary dropout with rate $p \in (0, 1)$ uses $
%   p_\xi
%     = \mathcal{B}\bigl(
%       \{0, \tfrac1{1-p}\}, 1-p
%     \bigr)
% $, whereas in Gaussian dropout $
%   \xi \sim \mathcal{N}(1, \alpha)
% $ for $\alpha = \tfrac{p}{1-p}$.
% This suggests making $\alpha$ a variational parameter
% and optimizing it in \eqref{eq:elbo_general} with an appropriate penalty term. Thus,
%
Variational Dropout with assumes mean field Gaussian approximation $
  q_\theta(\omega)
    = \prod_j \mathcal{N}(\mu_j, \alpha_j \mu_j^2)
$ and factorized log-uniform prior $\pi(\omega)$ with $
  \pi(\omega_j) \propto \lvert \omega_j \rvert^{-1}
$. This unravels the divergence term in \eqref{eq:elbo} into a sum of the following
analytically intractable terms:
\begin{equation}  \label{eq:improper-kl-div-real}
  KL(q_\theta(\omega) \| \pi(\omega))
    \propto
      \frac12 \mathbb{E}_{\varepsilon \sim \mathcal{N}(0, 1)}
        \log{\Bigl\lvert \tfrac1{\sqrt{\alpha}} + \varepsilon \Bigr\rvert^2}
      % \frac12 \mathbb{E}_{\varepsilon \sim \mathcal{N}(0, 1)}
      %   \log{\bigl\lvert 1 + \sqrt{\alpha} \varepsilon \bigr\rvert^2}
      % - \frac12 \log{\alpha}
  \,.
\end{equation}
\cite{kingma_variational_2015} approximate this expression by a non-linear polynomial
regression for \verify{truncated} $\alpha \in (0, 1)$. This approximation was refined in
\cite{molchanov_variational_2017}, both in terms of the domain of $\alpha$ and accuracy. The
paper considered regression with sigmoid and soft-plus basis functions fit to the Monte-Carlo
estimator of \eqref{eq:improper-kl-div-real} for $\alpha$ varying over a fine logarithmic grid.
In sec.~\ref{sub:real-chisq-grad} we verify the approximation and confirm its accuracy, by
comparing its derivative to the exact derivative of \eqref{eq:improper-kl-div-real} with
respect to $\log \alpha$.
% (!) good for SGVB: gradient-based methods require unbiased gradient estimates

Flexibility in specification of $q_\theta$ enables structured dropout with shared
$\alpha$ across groups of parameters $\omega$. \cite{molchanov_variational_2017} study
Variational Dropout from a model sparsification perspective, optimizing $\alpha$ for each
individual parameter. Another variant of Bayesian dropout, called \textit{Automatic Relevance
Determination}, is developed in \cite{kharitonov_variational_2018}. The key difference is
that the log-uniform factor $\pi(\omega_{ij})$ is replaced by a proper Gaussian prior $
  \mathcal{N}(0, \tau^{-1}_{ij})
$ with learnable precision $\tau_{ij} > 0$. This method, known as \textit{Empirical Bayes}
\cite{citation_needed}, fits a prior distribution while performing variational inference.
Maximizing \eqref{eq:elbo} over $\lambda$, holding other parameters fixed, yields $
  \tau^* = {(\mu^2 + \sigma^2)}^{-1}
$, whence
\begin{equation}  \label{eq:ard-kl-div-real}
  KL(q_\theta(\omega) \| \pi(\omega))
    = \log{\bigl(1 + \tfrac1{\alpha} \bigr)}
    \,.
\end{equation}

Another contribution of \cite{kingma_variational_2015} is the analysis of the effects of the
\textit{local reparameterization trick} on the variance of the gradient of \eqref{eq:elbo}.
This trick, proposed in \cite{wang_fast_2013} to speed up Gaussian dropout, utilizes the
closure of Gaussians under affine transformations, thereby translating uncertainty from the
global parameter noise to local noise in intermediate outputs within the network.
%
The stochastic output of a linear layer $
  y = W^\top x + b
$ with $
  W \in \mathbb{R}^{n\times m}
$ and independent $
  W_{ij} \sim \mathcal{N}(\mu_{ij}, \alpha_{ij} \mu_{ij}^2)
$ is equivalent in distribution to
\begin{equation}  \label{eq:r-gauss-trick}
    y \sim \prod_i \mathcal{N}\Bigl(
          \sum_j \mu_{ij} x_j + b_i,
          \sum_j \alpha_{ij} \mu_{ij}^2 x_j^2
      \Bigr)
    \,.
\end{equation}
%
\cite{kingma_variational_2015} show, that for the case of fully factorized Gaussian
posterior approximations, the trick decorrelates ELBO estimates within the minibatch,
making the gradient estimator $\nabla_\theta \tilde{\mathcal{L}}$ more statistically
efficient. \cite{molchanov_variational_2017} propose an improvement by introducing
\textit{additive noise parameterization}, which is relevant to model sparsification.
The paper argues that $(\mu, \alpha)$ parameterization should be abandoned in favour
of an equivalent $(\mu, \sigma^2)$ with the score $\alpha$ calculated using $
  \tfrac{\sigma^2}{\mu^2}
$ when needed. This makes $
  \nabla_\mu \tilde{\mathcal{L}}
$, which is shown in \cite{molchanov_variational_2017} to be independent of the noise,
injected by the trick \eqref{eq:r-gauss-trick}, and further stabilizes gradients in SGVB
\eqref{eq:elbo}.

@@

The study \cite{trabelsi_deep_2017} outlines building blocks for deep $\cplx$-valued
networks and describes suitable representation and operations including convolutional
and dense layers, $\cplx$-valued activations, complex batch-normalization and suitable
weight initialization. Their approach makes a complex-valued network into an intricately
connected $\real$-valued computational graph that respects $\cplx$-algebra.

In particular, linear layers
$
  L \colon \cplx^{\mathrm{[in]}}
    \to \cplx^{\mathrm{[out]}}
$ act upon their inputs as follows:
\begin{equation}  \label{eq:cplx-lin-op}
  \real^{2 \times \mathrm{[in]}}
    \to \real^{2 \times \mathrm{[out]}}
    \colon (x_r, x_i)
      \mapsto \Bigl(
        L_r x_r - L_i x_i,
        L_r x_i + L_i x_r
      \Bigr)
    \,,
\end{equation}
where $
  L_r, L_i
    \colon \real^{\mathrm{[in]}}
      \to \real^{\mathrm{[out]}}
$ are unique operators that make up $L = L_r + j L_i$. A $\cplx$-convolutional layer with
kernel $W$ can be implemented as two $\real$-convolutions with kernels $\Re{W}$ and $\Im{W}$
using \eqref{eq:cplx-lin-op}. Algebraic operations, trigonometric functions, are implemented
through operations in $\real^2$. Activations, considered in \cite{trabelsi_deep_2017} include
independent componentwise application of $\real$-valued non-linearities, $
  z = u + j v
    \mapsto \sigma(u) + j \sigma(v)
$, $\cplx$-differentiable exponential and trigonometric activations, and \textit{modRelu}
activation, which applies soft-thresholding to the modulus and preserves the angle: $
  z = r e^{j \phi}
    \mapsto e^{j \phi} (r - b)_+
  % z \mapsto z \bigl(
  %     1 - \tfracb{\lvert z \rvert}
  %   \bigr)_+
$ with a learnable threshold parameter $b \in \real$.

>>>>

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{natbib}
% \setcitestyle{authoryear,open={((},close={))}}

@@

Deep neural networks are an integral part of machine learning and data science toolset
for practical data-driven problem solving. Despite recent advances in generalization
and optimization theory specific to deep networks, deploying in actual embedded hardware
remains a challenge. Solving the storage and compute constraints encourages development
of model compression and sparsification methods, that focus on favourable tradeoff between
performance and size. Most popular techniques include weight pruning \citep{citation_needed},
model quantization from IEEE754 floating point to integer-based arithmetic \citep{citation_needed},
and dropout. The key assumption is that the more explicitly zero parameters there is,
the less computations is needed.

Bayesian Inference is a principled framework of reasoning about uncertainty. The core principle
is updating prior beliefs in accordance with the likelihood of empirical observations into
posterior belief. Under suitable modelling and prior assumptions, Bayesian methods can be
used towards model sparsification, \citep{kingma_variational_2015,molchanov_variational_2017},
and \citep{kharitonov_variational_2018}.

% complex nets and related stuff
data that is naturally complex-valued
studies have shown a potentially richer representation of complex nets
\citep{hirose2003,fijulamin2009}
% [3] Akira Hirose, Complex-valued neural networks: theories and applications, vol. 5, World Scientific, 2003.
% [4] Md Faijul Amin and Kazuyuki Murase, “Single-layered complex-valued neural network for real-valued classification problems,” Neurocomputing, vol. 72, no. 4-6, pp. 945–955, 2009.

@@

\citep{yang_complex_2019} develop complex transformer, with $\cplx$-attention and complex
encoder-decoder, and apply it to music transcription and wireless signal classification.

@@

papers related to complex-valued distributions \citep{pav_moments_2015,taubock_complex-valued_2012},
and \citep{karseras_caution:_2014}

Application of complex-valued networks is not new, \citep{citation_needed}. More recently
they were applied to music transcription and speech recognition \citep{trabelsi_deep_2017},
and \citep{yang_complex_2019}. The key motivation behind adoption of $\cplx$-networks is
pervasiveness of $\cplx$-valued data in digital signal processing \citep{citation_needed},
acoustic and radio signal analysis \citep{citation_needed}, and quantum systems modelling
\citep{citation_needed}, etc. All these applications require compliance with algebraic
operations in the field of complex numbers.

@@

The posterior itself is computationally intractable, save for the relatively simple cases
\citep{citation_needed}, and therefore exact inference is traded for tractability and scalability
of Variational Inference. Instead of deriving $p(\omega \mid D)$ using the rule, this approach
recasts posterior inference into variational optimization problem: finding $q$ that is close
to $p(\omega \mid D)$ in terms of a proximity score $\rho$
\begin{equation}  \label{eq:variational-progam}
  q_*
    \in \arg \min_{\theta} \rho\bigl(
      q_\theta(\omega), p(\omega \mid D)
    \bigr)
    \,,
\end{equation}
over a tractable parametric family of distributions on the parameter space $
  \mathcal{Q} = \{Q_\theta(d\omega) \colon \theta \in \Theta\}
$, $\theta$ -- generic variational parameter. The class $\mathcal{Q}$ is picked so that
its members have tractable densities $
  q_\theta(\omega) d\omega
$, can be sampled from and possess tractable $\log$-derivatives $
  \omega \mapsto \nabla_\theta \log q_\theta(\omega)
$, \citep{williams_simple_1992}, or can be represented as differentiable transformations
of non-parametric noise, \citep{kingma_auto-encoding_2014,figurnov_implicit_2019}.
% (reparameterization trick, pathwise gradieent) $q_{\theta}(d\omega)$ is a push-forward
% of $p(d\varepsilon)$ by a differentiable map $g(\varepsilon; \theta)$.
% % http://stillbreeze.github.io/REINFORCE-vs-Reparameterization-trick/#fn:1

Kullback-Leibler divergence of density $q$ from $p$ is
\begin{equation}  \label{eq:kl-div-def}
  KL(q \| p)
    % = \int \frac{dQ}{dP} \log{\frac{dQ}{dP}} P(d\omega)
    % = \int \frac{q(\omega)}{p(\omega)} \log{\frac{q(\omega)}{p(\omega)}} p(\omega) d\omega
    % = \mathbb{E}_{\omega \sim q}
    %   \log\frac{q(\omega)}{p(\omega)}
    = \mathbb{E}_{\omega \sim q}
      \log{q(\omega)}
    - \mathbb{E}_{\omega \sim q}
      \log{p(\omega)}
    \,,
\end{equation}
% and, despite it not being a metric and requiring distributions with nested support,
and is the natural choice for $\rho$, the main reason being that it satisfies the following
identity for well-behaved densities:
\begin{equation}  \label{eq:kl-div-master}
  \log p(D)
    % = \mathbb{E}_{q(\omega)} \log \frac{p(D, \omega)}{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{q(\omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    % - \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{\pi(\omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
    = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    - KL(q(\omega) \| \pi(\omega))
    + KL(q(\omega) \| p(\omega \mid D))
    \,.
\end{equation}
This identity yields an equivalent problem for \eqref{eq:variational-progam}:
maximizing the \textit{evidence lower bound} (ELBO)
\begin{equation}  \label{eq:elbo_general}
\begin{aligned}
  & \underset{\theta, \phi, \lambda}{\text{maximize}}
    & \mathcal{L}(\theta; \phi, \lambda)
      & = \mathbb{E}_{\omega \sim q_{\theta}}
          \log p_{\phi}(D \mid \omega)
        - KL(q_{\theta}(\omega) \| \pi_{\lambda}(\omega))
        % \mathbb{E}_{\omega \sim q_{\theta}}
          % \log\frac{q_\theta(\omega)}{\pi_{\lambda}(\omega)}
  % \\
  % & &
  %     & = \mathbb{E}_{\omega \sim q_{\theta}}
  %         \log{p_{\phi}(D \mid \omega) \pi_{\lambda}(\omega)}
  %       + \mathbb{H}(q_{\theta})
  \,,
\end{aligned}
\end{equation}
where the prior and likelihood are allowed to depend on explicit parameters. The most
commonly used algorithm to optimize \eqref{eq:elbo_general} is EM-algorithm. Other
variational objectives are possible, provided $p(\omega \mid D)$ is probed only through
$\log p(D \mid \omega)$ and $\log \pi(\omega)$, \citep{ranganath_operator_2018}.

\textit{Stochastic Gradient Variational Bayes} (SGVB), proposed by \citep{kingma_auto-encoding_2014},
is a differentiable unbiased Monte-Carlo estimator of \eqref{eq:elbo_general}, which makes
it possible to employ stochastic gradient-based optimization methods. If $
  \omega \sim q_{\theta}
$ is equivalent in distribution to $
  \omega = g(\varepsilon; \theta)
$ for some non-parametric random variable $
    \varepsilon \sim p_\varepsilon
$ and differentiable $
  g(\varepsilon; \theta)
$, then the SGVB is
\begin{equation}  \label{eq:sgvb}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta}(\omega) \| \pi_{\lambda}(\omega))
      % + \frac{N}{\lvert B \rvert} \sum_{i\in B}
        % \log p(z_i \mid \omega_i)b
      + \frac1{L} \sum_{l=1}^L
        \log p_{\phi}(D \mid \omega_{il})
          \Big\vert_{\omega_{l} = g(\varepsilon_{l}; \theta)}
      % =
      % - KL(q_{\theta}(\omega) \| \pi(\omega))
      % + N \mathbb{E}_{B \sim \{D\}_M \otimes q_\theta^M}
        % \hat{\mathbb{E}}_{z,\omega \sim B}
          % \log p(z \mid \omega)
    \,, 
\end{equation}
where $
  (\varepsilon_{l})_{l=1}^L
$ is an iid sample from $p_\varepsilon$. Intractable KL-divergence terms can be replaced
by differentiable unbiased Monte-Carlo estimators, \citep{kingma_auto-encoding_2014}.
%
For the purposes of this study, we assume that the observed data is independent, and
use minibatch version of \eqref{eq:sgvb} with $L=1$:
\begin{equation}  \label{eq:elbo}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta}(\omega) \| \pi_{\lambda}(\omega))
      + \frac{N}{M} \sum_{i=1}^M
        \log p_{\phi}(z_i \mid g(\varepsilon_i; \theta))
    \,,
\end{equation}
for a random subsample $(z_i)_{i=1}^M$ from $D$ and iid $
  (\varepsilon_i)_{i=1}^M \sim p_\varepsilon
$.
%
\todo{DSVI}

@@

With special family of posterior approximation, variational inference can be used as
a regularization method and as model sparsification technique. \citet{kingma_variational_2015}
consider Dropout \citep{hinton_improving_2012}, DropConnect, \citep{wan_regularization_2013},
and Gaussian dropout, \citep{srivastava_dropout_2014,wang_fast_2013} through the lens
of Bayesian inference methods and proposes \textit{Variational Dropout}. They argue that
the multiplicative noise introduced by these methods induces a distribution equivalent to
a fully factorized variational posterior of the form $
  q_\theta(\omega) = \prod_j q_{\theta}(\omega_j)
$, where $q_{\theta}(\omega_j)$ is $\omega_j = \mu_j \xi_j$ with $
  \xi_j \sim p_\theta(\xi_j)
$ iid from some $p_\theta(\xi)$.
%
% Binary dropout with rate $p \in (0, 1)$ uses $
%   p_\xi
%     = \mathcal{B}\bigl(
%       \{0, \tfrac1{1-p}\}, 1-p
%     \bigr)
% $, whereas in Gaussian dropout $
%   \xi \sim \mathcal{N}(1, \alpha)
% $ for $\alpha = \tfrac{p}{1-p}$.
% This suggests making $\alpha$ a variational parameter
% and optimizing it in \eqref{eq:elbo_general} with an appropriate penalty term. Thus,
%
Variational Dropout assumes fully factorized Gaussian approximate posterior $
  q_\theta(\omega)
    = \prod_j \mathcal{N}(\mu_j, \alpha_j \mu_j^2)
$ and factorized log-uniform prior $\pi(\omega)$ with $
  \pi(\omega_j) \propto \lvert \omega_j \rvert^{-1}
$. This unravels the divergence term in \eqref{eq:elbo} into a sum of the following
analytically intractable terms:
\begin{equation}  \label{eq:improper-kl-div-real}
  KL(q_\theta(\omega) \| \pi(\omega))
    \propto
      \frac12 \mathbb{E}_{\varepsilon \sim \mathcal{N}(0, 1)}
        \log{\Bigl\lvert \tfrac1{\sqrt{\alpha}} + \varepsilon \Bigr\rvert^2}
      % \frac12 \mathbb{E}_{\varepsilon \sim \mathcal{N}(0, 1)}
      %   \log{\bigl\lvert 1 + \sqrt{\alpha} \varepsilon \bigr\rvert^2}
      % - \frac12 \log{\alpha}
  \,.
\end{equation}
\citet{kingma_variational_2015} approximate this expression by a non-linear polynomial
regression for \verify{truncated} $\alpha \in (0, 1)$. \citet{molchanov_variational_2017}
refine the approximation by using sigmoid and soft-plus basis functions.
%
In sec.~\ref{sub:real-chisq-grad} we verify the approximation against Monte-Carlo
estimates of \eqref{eq:improper-kl-div-real} for $\alpha$ varying over a fine logarithmic
grid, and confirm its accuracy, by comparing its derivative to the exact derivative of
\eqref{eq:improper-kl-div-real} with respect to $\log \alpha$.
% (!) good for SGVB: gradient-based methods require unbiased gradient estimates

Flexibility in specification of $q_\theta$ enables structured dropout with shared
$\alpha$ across groups of parameters $\omega$. \citet{molchanov_variational_2017} study
Variational Dropout from a model sparsification perspective, optimizing $\alpha$ for each
individual parameter. Another variant of Bayesian dropout, called \textit{Automatic Relevance
Determination}, is developed in \citep{kharitonov_variational_2018}. The key difference is
that the log-uniform factor $\pi(\omega_{ij})$ is replaced by a proper Gaussian prior $
  \mathcal{N}(0, \tau^{-1}_{ij})
$ with learnable precision $\tau_{ij} > 0$. This method, known as \textit{Empirical Bayes}
\citep{citation_needed}, fits a prior distribution while performing variational inference.
Maximizing \eqref{eq:elbo} over $\lambda$, holding other parameters fixed, yields $
  \tau^* = {(\mu^2 + \sigma^2)}^{-1}
$, whence
\begin{equation}  \label{eq:ard-kl-div-real}
  KL(q_\theta(\omega) \| \pi(\omega))
    = \log{\bigl(1 + \tfrac1{\alpha} \bigr)}
    \,.
\end{equation}

Another contribution of \citep{kingma_variational_2015} is the analysis of the effects of the
\textit{local reparameterization trick} on the variance of the gradient of \eqref{eq:elbo}.
This trick, proposed in \citep{wang_fast_2013} to speed up Gaussian dropout, utilizes the
closure of Gaussians under affine transformations, thereby translating uncertainty from the
global parameter noise to local noise in intermediate outputs within the network.
%
The stochastic output of a linear layer $
  y = W^\top x + b
$ with $
  W \in \mathbb{R}^{n\times m}
$ and independent $
  W_{ij} \sim \mathcal{N}(\mu_{ij}, \alpha_{ij} \mu_{ij}^2)
$ is equivalent in distribution to
\begin{equation}  \label{eq:r-gauss-trick}
    y \sim \prod_i \mathcal{N}\Bigl(
          \sum_j \mu_{ij} x_j + b_i,
          \sum_j \alpha_{ij} \mu_{ij}^2 x_j^2
      \Bigr)
    \,.
\end{equation}
%
\citet{kingma_variational_2015} show, that for the case of fully factorized Gaussian
posterior approximations, the trick decorrelates ELBO estimates within the minibatch,
making the gradient estimator $\nabla_\theta \tilde{\mathcal{L}}$ more statistically
efficient. \citet{molchanov_variational_2017} propose an improvement by introducing
\textit{additive noise parameterization}, which is relevant to model sparsification.
They argue that $(\mu, \alpha)$ parameterization should be abandoned in favour of an
equivalent $(\mu, \sigma^2)$ with the score $\alpha$ calculated using $
  \tfrac{\sigma^2}{\mu^2}
$ when needed. This makes $
  \nabla_\mu \tilde{\mathcal{L}}
$, which is shown in \citep{molchanov_variational_2017} to be independent of the noise,
injected by the trick \eqref{eq:r-gauss-trick}, and further stabilizes gradients in SGVB
\eqref{eq:elbo}.

@@

\citet{trabelsi_deep_2017} outline the building blocks for deep $\cplx$-valued networks
and describe suitable representation and operations including convolutional and dense layers,
$\cplx$-valued activations, complex batch-normalization and suitable weight initialization.
Their approach makes a complex-valued network into an intricately connected $\real$-valued
computational graph that respects $\cplx$-algebra.

In particular, linear layers
$
  L \colon \cplx^{\mathrm{[in]}}
    \to \cplx^{\mathrm{[out]}}
$ act upon their inputs as follows:
\begin{equation}  \label{eq:cplx-lin-op}
  \real^{2 \times \mathrm{[in]}}
    \to \real^{2 \times \mathrm{[out]}}
    \colon (x_r, x_i)
      \mapsto \Bigl(
        L_r x_r - L_i x_i,
        L_r x_i + L_i x_r
      \Bigr)
    \,,
\end{equation}
where $
  L_r, L_i
    \colon \real^{\mathrm{[in]}}
      \to \real^{\mathrm{[out]}}
$ are unique operators that make up $L = L_r + j L_i$. A $\cplx$-convolutional layer with
kernel $W$ can be implemented as two $\real$-convolutions with kernels $\Re{W}$ and $\Im{W}$
using \eqref{eq:cplx-lin-op}. Algebraic operations, trigonometric functions, are implemented
through operations in $\real^2$. Activations, considered in \citep{trabelsi_deep_2017} include
independent componentwise application of $\real$-valued non-linearities, $
  z = u + j v
    \mapsto \sigma(u) + j \sigma(v)
$, $\cplx$-differentiable exponential and trigonometric activations, and \textit{modRelu}
activation, which applies soft-thresholding to the modulus and preserves the angle: $
  z = r e^{j \phi}
    \mapsto e^{j \phi} (r - b)_+
  % z \mapsto z \bigl(
  %     1 - \tfracb{\lvert z \rvert}
  %   \bigr)_+
$ with a learnable threshold parameter $b \in \real$.

Complex networks are, in general, non-holomorphic, i.e. not $\cplx$-differentiable, which is
exacerbated by the loss being real-valued. Indeed a $\cplx$-differentiable real-valued function is
necessarily trivial by Cauchy-Riemann conditions. This issue is dealt with by employing Wirtinger,
or $\cplx\real$ calculus, which generalizes holomorphic calculus to non-holomorphic functions
of $\cplx$-argument, \citep{adali_complex-valued_2011}. The rules it defines satisfy intuitive
linearity, product and chain rules, and respect complex conjugation, but most importantly the
$\cplx\real$-differential of a function on $\cplx$ coincides with its classical differential as
a function on $\real^2$. Essentially, it allows straightforward retrofiting of $\cplx$-valued
networks into existing $\real$ deep learning auto-differentiation frameworks. It was considered
as a basis for $\cplx$ version of backpropagation in \citep{benvenuto_complex_1992} and
\citep{trabelsi_deep_2017}, (see sec.~\ref{sub:wirtinger_calculus} for details).
%
\todo{read on holomorphic nets}
